---
title: "DS202 Lab 3"
author: "Muhammad Khairi Norizan"
date: "10/3/2021"
output: html_document
---

```{r, echo=FALSE}
dat <- readxl::read_xls('./GSS.xls')

library(dplyr)
library(ggplot2)


# Data Engineering
dat <- dat %>% rename(Year = `Gss year for this respondent`, Happiness = `General happiness`, Politics = `Political party affiliation`, FamilyIncome = `Total family income`, Gender = `Respondents sex`, YearsSchoolCompleted = `Highest year of school completed`, MaritalStatus = `Marital status`, HoursWorkPerWeek = `Number of hours usually work a week`, id = `Respondent id number`, Ballot = `Ballot used for interview`)

# Remove not numeric and NA rows
dat <- dat %>% 
  filter(!is.na(as.numeric(YearsSchoolCompleted)))

# Convert column years school completed to numerical values
dat$YearsSchoolCompleted = strtoi(dat$YearsSchoolCompleted)
```

## Exploration

#### Question 1

Find the average years of education for the respondents with each marital status. Arrange your output in a meaningful order and print. Describe any patterns you find.

```{r}
selectedDat <- dat %>% 
  group_by(MaritalStatus) %>% 
  summarize(AvgYearsSchoolCompleted = mean(YearsSchoolCompleted))

selectedDat <- selectedDat %>% 
  arrange(desc(AvgYearsSchoolCompleted))

print(selectedDat)

ggplot(selectedDat, aes(x=MaritalStatus, y=AvgYearsSchoolCompleted, fill=MaritalStatus)) + geom_bar(stat="identity")
```

#### Question 2

Create a single data frame containing records for the better educated respondents with each marital status. A "better educated" respondent is someone who has strictly more years of education than the average among those with the same marital status. Print the structure of the data frame.

```{r,}
betterEducatedFilter <- dat %>% 
  group_by(MaritalStatus) %>% 
  mutate(AvgYearsSchoolCompleted = mean(YearsSchoolCompleted, na.rm = TRUE)) %>% 
  filter(YearsSchoolCompleted >= AvgYearsSchoolCompleted)

print(str(betterEducatedFilter))

```

#### Question 3

How is the happiness of a respondent related to his/her marriage status? Define that a person is happy if the response to question is "Very happy" or "Pretty happy".

```{r}
happiness <- dat %>% 
  filter(Happiness %in% c('Very happy', 'Pretty happy', 'Not too happy')) %>%
  mutate(HappyOrNot = (Happiness %in% c('Very happy', 'Pretty happy')))

ggplot(data = happiness) + geom_bar(mapping = aes(x = MaritalStatus, fill = HappyOrNot), position = 'dodge')

```
\
The graph above shows that people are mostly happy despite their marital status. Obviously, "Married" group has the most number of happy people compared to the other groups. "Never married" group follow as the second group that has the most number of happy people, and "Divorced", "Widowed", "Separated" follow after that. 
\
\
Despite "Married" group has the most number of happy people, the same group also has the most count for not happy numbers. Similar to the number of happy count, "Never married" follow as the close second that has the most count for not happy. 

#### Question 4

Does party affiliation affect the relationship you found in the last question

```{r}
filteredHappiness <- happiness %>% filter(Politics %in% c('Independent', 'Strong democrat', 'Strong republican', 'Other party'))

ggplot(data = filteredHappiness) + geom_bar(mapping = aes(x = MaritalStatus, fill = HappyOrNot), position = 'dodge') + facet_wrap(~Politics) + coord_flip()
```

\
I don't think its appropriate to say politic party affiliation affect the relationship. However, looking at each facet (represent party affiliation), every politic party affiliation group (Independent, Other party, Strong democrat, Strong republican) are usually happy when they are married. 
\
\
However, the numbers of count are different for each politic party affiliation group. "Strong democrat" has the most number of happy married people compared to the other 3 groups, followed by "Independent", "Strong republican", and "Other party" respectively. 

#### Question 5

Explore an interesting question (not covered by the previous questions), and answer the question using the GSS data.
\
**Question: How Gender and MaritalStatus affecting the Year School Completed**

```{r}
selectedDat <- dat %>% 
  group_by(MaritalStatus, Gender) %>% 
  summarize(AvgYearsSchoolCompleted = mean(YearsSchoolCompleted))

ggplot(selectedDat, aes(x=MaritalStatus, y=AvgYearsSchoolCompleted, fill=Gender)) + geom_bar(stat="identity", position = 'dodge')
```
\
From the above graph, we can see that for most of the marital status, male gender would usually have an equal or better average years school completed than female gender. Differently, in the case of "Never Married" marital status, the female gender has slightly higher average years school completed than the male gender. 
\
\
According to the graph, apparently, "No answer" group has the highest average years school completed. Second highest is "Never Married", and followed by "Divorced", "Married", "Separated", and "Widowed".